Seminar:\\ Quantitative Methoden

### 2024-02-15 draft2.tex
validity

    %  \noindent % otherwise the line will be too wide by \parindent
    \begin{tabularx}{\textwidth}{lXlr}

@@

    %  \noindent % otherwise the line will be too wide by \parindent
    \begin{tabularx}{\textwidth}{lXlr}

    \hline & \\[-1.0em]
    \multicolumn{2}{l}{Probability of having more than one child}  & Probability of being married &  \\
    \hline & \\[-1.0em]
        First-born boy & -0.013***  & First-born boy  & 0.001   \\
        St. err.       & 0.004      & St. err.        & 0.001   \\
        \\[-1.0em]                                                     
        Baseline Girl: & 0.818      & Baseline Girl:  & 0.985   \\
        St. err.       & 0.038      & St. err.        & 0.001   \\
        Percent effect & -1.589***  & Percent effect  & 0.082   \\
        No. obs.       & 39,901     & No. obs.        & 39,901 \\
    \hline
    \end{tabularx}

@@

\section{Main Summary of Paper}
% • What is the research question(s)?
% • What is the main contribution?

@@

% No further minimum wage increases can lead to deterioration of the reduction
%Discussion of Employment effect
% No dummy for Treatment group is used, doesn't show how far groups are apart, and which one has a larger gender pay gap.

% External validity
% Germany as an extreme case
%Other countries

@@

@@

Another interesting paper about the minimum wage effect on the gender wage gap is written by \citet{Robinson05UK}, studying the effect the national minimum wage in the UK has on the gender pay gap. The paper creates a theoretical model and discusses the prerequisites for the effect of the minimum wage on the gender pay gap. 
%rewrite sentence ->
Not only do Women have to be disproportionally affected by the minimum wage but the average wage difference between the wage payed before and the minimum wage has to be also larger
% <- rewrite sentence
\Citeauthor{Robinson05UK} uses a similar approach, looking at regions where the fraction of women earning under the minimum wage compared to men working under the minimum wage is large. The result of this difference-in-differences estimation is (although her coefficients are mainly insignificant) that the minimum wage led to a decrease relative decrease of the gender pay gap in the treatment regions.
Despite this similar result to Germany there is a big difference here.
Because men were overall even lower paid than women, the introduction of the minimum wage increased the gender pay gap. This shows that despite finding that the minimum wage reduces the gender wage gap in regions where women are affected relative to the other regions, the overall effect might differ.

\section{Connection to empirical project} %maybe only half a page, additional empirical discussion if time
% Interesting future research with panel data effect on mothers
%Maybe quote child penalty atlas
%Future

>>>>

    %  \noindent % otherwise the line will be too wide by \parindent
    \begin{tabularx}{\textwidth}{lXlc}

@@

    %  \noindent % otherwise the line will be too wide by \parindent
    \begin{tabularx}{\textwidth}{lXlc}

    \hline & \\[-1.0em]
    \multicolumn{2}{l}{Probability of having more than one child}  & \multicolumn{2}{l}{Probability of being married} \\
    \hline & \\[-1.0em]
        First-born boy & -0.013***  & First-born boy  & 0.001   \\
        St. err.       & 0.004      & St. err.        & 0.001   \\
        \\[-1.0em]                                                     
        Baseline Girl: & 0.818      & Baseline Girl:  & 0.985   \\
        St. err.       & 0.038      & St. err.        & 0.001   \\
        Percent effect & -1.589***  & Percent effect  & 0.082   \\
        No. obs.       & 39,901     & No. obs.        & 39,901 \\
    \hline
    \end{tabularx}

@@

\section{Main Summary of Paper}
% • What is the research question(s)?
% • What is the main contribution?
\subsection{Overview} \label{overview}

@@

% No further minimum wage increases can lead to deterioration of the reduction
%Discussion of Employment effect
Another big issue the possible influence of the employment effect. 
If the minimum wage leads to job losses for low-paid workers, and women are disproportional overrepresented in that group, job losses decrease the minimum wage gap. But this effect is unwanted, since it doesn't reduce the wage gap by raising women's wages, but by low-income women becoming unemployed and thus raising the average women's wage. 
This possible chanel of reducing the gender wage gap is now further analyzed by \citet{CALIENDO22} citing that there are, if any, only small employment effects. But one of the cited papers for this, \citep{Caliendo2017short} adds that instead of unemployment, a reduction of working hours might be used.
If employees have to reduce their working hours in response to the minimum wage, employees might get payed more per hour, but still earn the same or less overall.
But this possible stagnation of overall income is not be accounted for her, since the gender pay gap is calculated by hourly wages.
These effects should be examined in the future, since they might lead to an overestimation of the results, and are mainly neglected in the paper.

% No dummy for Treatment group is used, doesn't show how far groups are apart, and which one has a larger gender pay gap.
Like in \Cref{overview} mentioned, Germany is an extreme case, with gender pay gap initially more than double that of the EU average for the for the first decile. While this means that the magnitude of the effect would be lower in other countries, the effect should be the same for most counties where women are overrepresented at the lower end of the wage distribution. To additionally validate this, the next section compares the results of similar studies in different countries.
% External validity
% Germany as an extreme case
%Other countries

@@

A interesting paper about the minimum wage effect on the gender wage gap is written by \citet{Robinson05UK}, studying the effect the national minimum wage in the UK has on the gender pay gap. The paper creates a theoretical model and discusses the prerequisites for the effect of the minimum wage on the gender pay gap. 
%rewrite sentence ->
Not only do Women have to be disproportionally affected by the minimum wage but the average wage difference between the wage payed before and the minimum wage has to be also larger
% <- rewrite sentence
\Citeauthor{Robinson05UK} uses a similar approach, looking at regions where the fraction of women earning under the minimum wage compared to men working under the minimum wage is large. The result of this difference-in-differences estimation is (although her coefficients are mainly insignificant) that the minimum wage led to a decrease relative decrease of the gender pay gap in the treatment regions.
Despite this similar result to Germany there is a big difference here.
Because men were overall even lower paid than women, the introduction of the minimum wage increased the gender pay gap. This shows that despite finding that the minimum wage reduces the gender wage gap in regions where women are affected relative to the other regions, the overall effect might differ.

@@

\section{Connection to empirical project} %maybe only half a page, additional empirical discussion if time
% Interesting future research with panel data effect on mothers
%Maybe quote child penalty atlas

%factors affecting women labor force participation and pay
%Future
